CONEM2024-0025\\ ABC MULTICRITERIA CLASSIFICATION WITH ANALYTICAL HIERARCHY PROCESS: HOW TO FORCE CONSISTENCY OF PAIRWISE COMPARISONS MATRIX

### 2024-02-17 accepted/CONEM2024/FCPCOM/CONEM2024.tex
work in progress - improving introduction

    \subsection{Pairwise Comparisons Matrix and Consistency Rate}

@@

In our study, we observed that one of the crucial points for correctly applying multicriteria ABC classification is correctly attributing the weights for the initial ABC classification.

>>>>

	As stated by \cite{Kiran2019}, "production planning and control (PP\&C) ... is responsible for ensuring the availability \emph{of} all materials, part of assembly at the right time, at the right place, and in \emph{the} right quantities to enable the progress of operations according to the predetermined schedules at the minimum possible costs." The author warns that factories producing large volumes of standardized products need uncomplicated production control techniques, but factories producing a variety of products and product mixes also need meticulous control with complex production control techniques. In the latter case, prior strategic planning becomes impressible for better use of the productive capacity and to avoid investment losses in unnecessary or unprofitable material and, given the wide variety of products, it's crucial to start any strategic planning by understanding the importance of all products marked by the campany.

	In this work, we present a study with three companies from three different sectors, which market a wide variety of products, in some cases, highly personalized products. The interest of this study is to identify the main products, that is, the products that company must prioritize, so that this information can guide the strategic planning. Also, we identify the need to consider several criteria for such a classification, therefore we used a multicriteria approach.

	The importance of multicriteria classification for inventory management has been widely emphasized in scientific literature and several multicriteria methods have been used, being ABC multicriteria classification with Analytical Hierarchy Process one of the most common techniques applied for this purpose \citep{Flores1992, AltayGuvenir1998, Lolli2014, Balaji2014}. But just a few recent works talk about multicriteria classification applied to strategic planning \citep[\emph{e.g.},][]{BarbosaDePaula2022, Pereira2023, MarianoRibeiro2023}, and the multicriteria methods applied are diverse. \citet{Odu2019} emphasize that, when applying some multicriteria method, it is important to pay particular attention to the objectivity factors of criteria weights. The author provides an important overview of different weighting methods applicable to multicriteria techniques.

	In the present work, we opted for the multicriteria ABC classification and adopted the Analytic Hierarchy Process (AHP) \citep{Saaty1987} to assign the weights to the different criteria. It is important to note that the Analytic Hierarchy Process derives the weights of the criteria based on the eigenvector of a pairwise comparison matrix which is constructed manually according to an empirical evaluation. According to \cite{Saaty1987}, this matrix must be antisymmetric and must have a consistency rate of less than 0.1.  The author reports that a software package supporting the AHP, called Expert Choice, for the IBM PC, was used in his work for making eigenvector calculations and guiding the decision maker to improve matrix inconsistency if needed. During the execution of the work we present in this paper, we found that there is a strong difficulty in the calculation of the eigenvector, as well as for the construction of consistent pairwise comparison matrices and we had no access to the Expert Choice software package, so we developed an open source COPSolver library for Analytic Hierarch Process and multicriteria ABC classification based on a simple algorithm for forcing the consistency of such matrices. We present this algorithm in this paper. The Eigen library from eigen.tuxfamily.org was coupled to COPSolver for eigenvectors calculation of the matrices. Also, we identified that the assignment of weights to ABC classification can be done using different criteria and we observed that such an assignment generates a bias in the multicriteria ABC classification. Therefore we also developed a new methodology for assigning weights in a balanced way for ABC classification and we proposed a new multicriteria weighting equation for the multicriteria classification COPSolver library. We also present both propositions in this paper and an in-depth analysis of the results found by the new library developed for COPSolver.

@@

As discussed earlier, in our study, we observed that one of the crucial points for correctly applying multicriteria ABC classification is correctly attributing the weights for the initial ABC classification.
